---
title: "mschart R package"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![R build status](https://github.com/ardata-fr/mschart/workflows/R-CMD-check/badge.svg)](https://github.com/ardata-fr/mschart/actions)
[![version](https://www.r-pkg.org/badges/version/mschart)](https://CRAN.R-project.org/package=mschart)
![cranlogs](https://cranlogs.r-pkg.org/badges/mschart)
![Active](http://www.repostatus.org/badges/latest/active.svg)

The `mschart` package provides a framework for easily creating charts for
'Microsoft PowerPoint' presentations, 'Microsoft Word' documents and
'Microsoft Excel' workbooks. It has to be used with package
[`officer`](https://davidgohel.github.io/officer/) that will produce the
charts in new or existing PowerPoint, Word or Excel files. With 'Microsoft
Charts', the data is integrated into the document and linked to the chart.
The result can be edited, annotated and resized. If the data is updated in
the document, the chart is also updated.

## Example

This is a basic example which shows you how to create a bar chart.

```{r example}
library(mschart)

sales <- data.frame(
  quarter = rep(c("Q1", "Q2", "Q3", "Q4"), each = 2),
  revenue = c(12, 9, 15, 11, 18, 14, 21, 17),
  region  = rep(c("EU", "US"), times = 4)
)

bars <- ms_barchart(
  data = sales, x = "quarter", y = "revenue", group = "region"
)
```

Then use package `officer` to send the object as a chart.

```{r eval=FALSE}
library(officer)
doc <- read_pptx()
doc <- add_slide(doc, layout = "Title and Content", master = "Office Theme")
doc <- ph_with(doc, value = bars, location = ph_location_fullsize())

print(doc, target = "example.pptx")
```

The same chart object can be dropped into an Excel sheet with
`officer::sheet_add_drawing()`. By default the chart's underlying data is
written next to the chart automatically.

You can also write the data first with `officer::sheet_write_data()` and ask
the chart to reuse it by passing `write_data = FALSE`. This is useful when
the data should appear at a specific position, or when several charts share
the same dataset:

```{r eval=FALSE}
wb <- read_xlsx()
wb <- add_sheet(wb, label = "sales")

# 1. Write the data on the sheet at the desired position.
wb <- sheet_write_data(wb, sheet = "sales", value = bars$data_series,
                       start_col = 1, start_row = 1)

# 2. Add the chart and tell it to reference the data already in place.
wb <- sheet_add_drawing(wb, sheet = "sales", value = bars,
                        write_data = FALSE,
                        start_col = 1, start_row = 1,
                        left = 5, top = 0.5, width = 6, height = 4)

print(wb, target = "example.xlsx")
```

At any moment, you can type `print(your_chart, preview = TRUE)` to preview the
chart in a temporary PowerPoint file. This requires a PowerPoint Viewer to be
installed on the machine.

## Installation

You can get the development version from GitHub:

```{r eval = FALSE}
devtools::install_github("ardata-fr/mschart")
```

Or the latest version on CRAN:

```{r eval = FALSE}
install.packages("mschart")
```


## Contributing to the package

### Bug reports

When you file a [bug report](https://github.com/ardata-fr/mschart/issues),
please spend some time making it easy for me to follow and reproduce. The more
time you spend on making the bug report coherent, the more time I can dedicate
to investigating the bug as opposed to the bug report.
