---
title: "Details"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r 'setup', include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)


dir.create("assets/pptx", recursive = TRUE, showWarnings = FALSE)
dir.create("assets/docx", recursive = TRUE, showWarnings = FALSE)
office_doc_link <- function(url){
  stopifnot(requireNamespace("htmltools", quietly = TRUE))
  htmltools::tags$p(  htmltools::tags$span("Download file "),
    htmltools::tags$a(basename(url), href = url), 
    htmltools::tags$span(" - view with"),
    htmltools::tags$a("office web viewer", target="_blank", 
      href = paste0("https://view.officeapps.live.com/op/view.aspx?src=", url)
      ), 
    style="text-align:center;font-style:italic;color:gray;"
    )
}

base_url <- "https://ardata-fr.github.io/mschart/articles/"
```

```{r}
library(mschart)
library(officer)
library(magrittr)
```


## Chart types

> In order to make to code shorter, we will use the following function to 
generate a PowerPoint document from a chart.

```{r}
gen_pptx <- function( chart, file ){
  doc <- read_pptx()
  doc <- add_slide(doc, layout = "Title and Content", master = "Office Theme")
  doc <- ph_with_chart(doc, chart = chart)
  print(doc, target = file)
  office_doc_link( url = paste0( "https://ardata-fr.github.io/mschart/articles/", file ) )
}
```

The following are the chart available in package `mschart`.

### Bar charts

#### grouped bars

```{r}
my_barchart_01 <- ms_barchart(
  data = browser_data, x = "browser",
  y = "value", group = "serie"
)
gen_pptx(my_barchart_01, file = "assets/pptx/gallery_bar_01.pptx")
```

#### vertical stacked bars

`as_bar_stack()` is an helper function. 

```{r}
my_barchart_02 <- as_bar_stack(my_barchart_01)
gen_pptx(my_barchart_02, file = "assets/pptx/gallery_bar_02.pptx")
```

#### horizontal stacked bars

* Values

```{r}
my_barchart_03 <- as_bar_stack(my_barchart_01, dir = "horizontal")
gen_pptx(my_barchart_03, file = "assets/pptx/gallery_bar_03.pptx")
```

* Percent 

```{r}
my_barchart_04 <- as_bar_stack(my_barchart_01, dir = "horizontal", 
  percent = TRUE)
gen_pptx(my_barchart_04, file = "assets/pptx/gallery_bar_04.pptx")
```

### Line charts


```{r}
data <- structure(
  list(
    supp = structure( c(1L, 1L, 1L, 2L, 2L, 2L), 
                      .Label = c("OJ", "VC"), class = "factor" ),
    dose = c(0.5, 1, 2, 0.5, 1, 2),
    length = c(13.23, 22.7, 26.06, 7.98, 16.77, 26.14)
  ),
  .Names = c("supp", "dose", "length"),
  class = "data.frame", row.names = c(NA,-6L)
)

lc_01 <- ms_linechart(data = data, x = "dose",
                      y = "length", group = "supp")
gen_pptx(lc_01, file = "assets/pptx/gallery_line_01.pptx")
```

#### With dates

Auto detection of dates is not yet implemented, sorry. You will need to specify the `num_fmt` argument. 

```{r}
lc_02 <- ms_linechart(data = browser_ts, x = "date",
                      y = "freq", group = "browser")
lc_02 <- chart_ax_x(lc_02, num_fmt = "m/d/yy")
gen_pptx(lc_02, file = "assets/pptx/gallery_line_02.pptx")
```

#### Stacked

```{r}
lc_03 <- chart_settings(lc_02, grouping = "percentStacked")
gen_pptx(lc_03, file = "assets/pptx/gallery_line_03.pptx")
```

### Area chats

```{r}
ac_01 <- ms_areachart(data = data, x = "dose",
                      y = "length", group = "supp")
gen_pptx(ac_01, file = "assets/pptx/gallery_area_01.pptx")

ac_02 <- ms_areachart(data = browser_ts, x = "date",
                      y = "freq", group = "browser")
ac_02 <- chart_ax_y(ac_02, cross_between = "between", num_fmt = "General")
ac_02 <- chart_ax_x(ac_02, cross_between = "midCat", num_fmt = "m/d/yy")
gen_pptx(ac_02, file = "assets/pptx/gallery_area_02.pptx")

ac_03 <- chart_settings(ac_02, grouping = "percentStacked")
gen_pptx(ac_03, file = "assets/pptx/gallery_area_03.pptx")
```

### Scatter plots

```{r}
sc_01 <- ms_scatterchart(data = mtcars, x = "disp",
                         y = "drat", group = "gear")
gen_pptx(sc_01, file = "assets/pptx/gallery_scatter_01.pptx")

sc_02 <- ms_scatterchart(data = mtcars, x = "disp",
                         y = "drat", group = "gear")
gen_pptx(sc_02, file = "assets/pptx/gallery_scatter_02.pptx")

```


## Chart settings

Chart settings are general options relative to the graphic layout, such as grouping options (stacked, clustered, etc.) or scatter plot style (markers only, lines and markers, etc.).

`chart_settings()` is a generic function, each type of chart has its own method. 

The following examples are illustrating existing chart types:


```{r}
my_bc_01 <- ms_barchart(
    data = browser_data, x = "browser", y = "value", group = "serie") %>% 
  chart_settings( dir = "vertical", grouping = "stacked", overlap = 100 )

my_bc_02 <- ms_barchart(
    data = browser_data, x = "browser", y = "value", group = "serie") %>% 
  chart_settings( dir = "vertical", grouping = "clustered", 
                  gap_width = 400, overlap = -100 )

my_sc_01 <- ms_scatterchart(
  data = mtcars, x = "disp", y = "drat") %>% 
  chart_settings(scatterstyle = "marker")

my_sc_02 <- ms_scatterchart(
  data = mtcars, x = "disp", y = "drat") %>% 
  chart_settings(scatterstyle = "lineMarker")
```

```{r results='hide'}
layout <- "Title and Content"
master <- "Office Theme"
read_pptx() %>% 
  add_slide(layout, master) %>% ph_with_chart(chart = my_bc_01) %>% 
  add_slide(layout, master) %>% ph_with_chart(chart = my_bc_02) %>% 
  add_slide(layout, master) %>% ph_with_chart(chart = my_sc_01) %>% 
  add_slide(layout, master) %>% ph_with_chart(chart = my_sc_02) %>% 
  print(target = "assets/pptx/chart_settings_01.pptx")
```

```{r echo=FALSE}
office_doc_link( url = paste0( base_url, "assets/pptx/theme_01.pptx" ) )
```

As bar charts can be complex to configure, helper function `barchart_options()` is provided.

## Theme

A theme is a set of parameters (i.e. colors, fonts) that can be applied to a chart. 
As for `ggplot` objects, a theme is controling the appearance of non-data 
components of the plot.

Let's first create a chart:

```{r}
my_bc <- ms_barchart(
  data = browser_data, x = "browser", y = "value", group = "serie") %>% 
  as_bar_stack( ) %>% 
  chart_labels(title = "Title example", xlab = "x label", ylab = "y label")
```

... and add that chart in a PowerPoint file with `officer`:

```{r results='hide'}
gen_pptx(my_bc, file = "assets/pptx/theme_00.pptx")
```

`mschart` is providing an interface to themes with function `set_theme()`. This function is 
using an object produced by `mschart_theme()`. The following are the supported arguments 
of the function:

```{r echo=FALSE}
do.call(htmltools::tags$ul, 
lapply( sort(formalArgs(mschart_theme)), 
        function( x ) {
          htmltools::tags$li(x)
        } ) )
```


```{r}
mytheme <- mschart_theme(
  axis_title = fp_text(color = "#0D6797", font.size = 20, italic = TRUE),
  main_title = fp_text(color = "#0D6797", font.size = 24, bold = TRUE),
  grid_major_line = fp_border(color = "#AA9961", style = "dashed"),
  axis_ticks = fp_border(color = "#AA9961") 
  )
my_bc <- set_theme(my_bc, mytheme)
gen_pptx(my_bc, file = "assets/pptx/theme_01.pptx")
```

Some elements are inheriting properties from other theme elements. For example, `axis_title_x` and 
`axis_title_y` inherit from `axis_title` and are all expecting a
[`fp_text`](https://davidgohel.github.io/officer/reference/fp_text.html); 
`grid_major_line_x` and `grid_major_line_y` inherit from `grid_major_line` and are 
all expecting a [`fp_border`](https://davidgohel.github.io/officer/reference/fp_border.html).


An helper function is provided to allow modifications of elements of the 
chart theme: 

```{r}
my_bc <- chart_theme( x = my_bc, 
  axis_title_x = fp_text(color = "red", font.size = 11),
  main_title = fp_text(color = "red", font.size = 15),
  legend_position = "t" )
gen_pptx(my_bc, file = "assets/pptx/theme_02.pptx")
```




