---
title: "Introduction to mschart"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r 'setup', echo = FALSE, message=FALSE, warning=FALSE}
dir.create("assets/pptx", recursive = TRUE, showWarnings = FALSE)
dir.create("assets/docx", recursive = TRUE, showWarnings = FALSE)
office_doc_link <- function(url){
  stopifnot(requireNamespace("htmltools", quietly = TRUE))
  htmltools::tags$p(  htmltools::tags$span("Download file "),
    htmltools::tags$a(basename(url), href = url), 
    htmltools::tags$span(" - view with"),
    htmltools::tags$a("office web viewer", target="_blank", 
      href = paste0("https://view.officeapps.live.com/op/view.aspx?src=", url)
      ), 
    style="text-align:center;font-style:italic;color:gray;"
    )
}
```

```{r}
library(mschart)
library(officer)
library(magrittr)
```

Package `mschart` lets R users to create Microsoft Office charts from data, 
and then add title, legends, and annotations to the chart object.

## Available charts

Names of these high level functions are all prefixed with `ms_`.
The following charts are the only available from all possible MS charts:

* barcharts: `ms_barchart()`
* linecharts: `ms_linechart()`
* scatter plots: `ms_scatterchart()`
* areacharts: `ms_areachart()`

*More chart types will be made available in future versions.*


## Quick start

First, we need to call an high level function (one of those starting with `ms_`). The following arguments will be required:

* `data`: the data.frame to be plotted
* `x`, `y` and eventually `group`: colnames for x and y axis and a grouping column

```{r}
my_barchart <- ms_barchart(data = browser_data, 
  x = "browser", y = "value", group = "serie")
```

### Write charts in Microsoft documents

Let's have a look at the result by sending it to a new PowerPoint presentation:

```{r results='hide'}
library(officer)
doc <- read_pptx()
doc <- add_slide(doc, layout = "Title and Content", master = "Office Theme")
doc <- ph_with_chart(doc, chart = my_barchart)
print(doc, target = "assets/pptx/barchart_01_stacked.pptx")
```

```{r echo=FALSE}
office_doc_link( url = paste0( "https://davidgohel.github.io/mschart/articles/", "assets/pptx/barchart_01_stacked.pptx" ) )
```

Or in a new Word document:

```{r results='hide'}
doc <- read_docx()
doc <- body_add_chart(doc, chart = my_barchart, style = "centered")
print(doc, target = "assets/docx/barchart_01_stacked.docx")
```

```{r echo=FALSE}
office_doc_link( url = paste0( "https://davidgohel.github.io/mschart/articles/", "assets/docx/barchart_01_stacked.docx" ) )
```


### Chart settings

Global chart settings can be modified with function `chart_settings`. Each type of chart has its own set of parameters. The following *stack* the bars for each group.

```{r}
my_barchart <- chart_settings( my_barchart, grouping = "stacked", gap_width = 50, overlap = 100 )
```

